---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
options(digits = 4, width = 160)
```

# hmatch: Tools for cleaning and matching hierarchically-structured data
<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/epicentre-msf/hmatch.svg?branch=master)](https://travis-ci.org/epicentre-msf/hmatch)
[![Codecov test coverage](https://codecov.io/gh/epicentre-msf/hmatch/branch/master/graph/badge.svg)](https://codecov.io/gh/epicentre-msf/hmatch?branch=master)
<!-- badges: end -->

An R package for cleaning and matching messy hierarchically-structured data
(e.g. country / region / district / municipality). The general goal is to match
sets of hierarchical values in a raw dataset to corresponding values within a
reference dataset, while accounting for potential discrepancies such as:

- variation in character case, punctuation, spacing, use of accents, or spelling
- variation in hierarchical resolution (e.g. some entries specified to municipality-level but others only to region)
- missing values at one or more hierarchical levels
- values entered at the wrong hierarchical level

### Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/hmatch")
```

### Matching strategies

##### Low-level

- **`hmatch`**: match hierarchical sequences up to the highest-resolution level specified within a given row of raw data, optionally allowing for missing values below the match level, and fuzzy matches (using the [stringdist](https://github.com/markvanderloo/stringdist) package)

##### Higher-level

- **`hmatch_tokens`**: match tokens rather than entire strings to allow for variation in multi-term names
- **`hmatch_permute`**: sequentially permute hierarchical columns to allow for values entered at the wrong level
- **`hmatch_parents`**: match values at a given hierarchical level based on shared sets of 'offspring'
- **`hmatch_settle`**: try matching at every level and settle for the highest-resolution match possible
- **`hmatch_manual`**: match using a user-supplied dictionary
- **`hmatch_split`**: implement any other `hmatch_` function separately at each hierarchical level, only on unique sequences
- **`hmatch_composite`**: implement a variety of matching strategies in sequence, from most to least strict

### String standardization

Independent of optional fuzzy matching with
[stringdist](https://github.com/markvanderloo/stringdist), `hmatch` functions
use behind-the-scenes string standardization to help account for variation in
character case, punctuation, spacing, or use of accents between the raw and
reference data. E.g.

```
              raw_value       reference_value  match
----------------------------------------------------
original:     ILE DE  FRANCE  Île-de-France    FALSE
standardized: ile_de_france   ile_de_france    TRUE
```

Users can choose default standardization (illustrated above), no
standardization, or supply their own preferred function to standardize strings
(e.g. `tolower`).

### Usage

#### Example dataset

The `hmatch` package contains example datasets `ne_raw` (messy geographical
data) and `ne_ref` (reference data derived from a shapefile), based on a small
subset of northeastern North America.

```{r}
library(hmatch)

head(ne_raw) # raw messy data

head(ne_ref) # reference data derived from shapefile
```


#### Example workflow

##### Basic hierarchical matching with `hmatch()`

We'll start with a simple call to `hmatch` to see which rows can be matched with
no extra magic.

```{r}
hmatch(ne_raw, ne_ref, pattern = "^adm")
```

There are still quite a few unmatched rows, and entry 'PID14' actually matches
two different rows within `ref`, so we'll press on. We can separate the matched
and unmatched rows using inner- and anti-joins respectively, specifically using
the "resolve_" join type here to only consider matches that are unique.

```{r}
(raw_match1 <- hmatch(ne_raw, ne_ref, pattern = "^adm", type = "resolve_inner"))

(raw_remain1 <- hmatch(ne_raw, ne_ref, pattern = "^adm", type = "resolve_anti"))
```

##### Fuzzy matching

Next we'll add in fuzzy-matching, using the default maximum string-distance of 1.

```{r}
hmatch(raw_remain1, ne_ref, pattern = "^adm", fuzzy = TRUE, type = "inner")
```

Only one additional *unique* match, so we'll again split and move on. Note that
we've been using the `pattern` argument above to specify the hierarchical
columns in `raw` and `ref`, but because the hierarchical columns have the same
names in `raw` and `ref` (and are the only matching column names), we can drop
the `pattern` argument for brevity.

```{r}
(raw_match2 <- hmatch(raw_remain1, ne_ref, fuzzy = TRUE, type = "resolve_inner"))

(raw_remain2 <- hmatch(raw_remain1, ne_ref, fuzzy = TRUE, type = "resolve_anti"))
```

##### Tokenized matching

Next let's try `hmatch_tokens`, which matches based on components of strings
(i.e. tokens) rather than entire strings.

```{r}
(raw_match3 <- hmatch_tokens(raw_remain2, ne_ref, type = "resolve_inner"))

(raw_remain3 <- hmatch_tokens(raw_remain2, ne_ref, type = "resolve_anti"))
```

##### Permutation matching

If there are any values entered at the wrong hierarchical level, we can try
systematically permuting the hierarchical columns before matching.

```{r}
(raw_match4 <- hmatch_permute(raw_remain3, ne_ref, type = "resolve_inner"))

(raw_remain4 <- hmatch_permute(raw_remain3, ne_ref, type = "resolve_anti"))
```


##### The toughest cases

For the remaining rows that we haven't yet matched, there a few options. We
could use `hmatch_settle()` to settle for matches below the highest-resolution
level specified within a given row of `raw`. We could also do some 'manual'
comparison of the raw and reference datasets and create a dictionary to recode
values within `raw` to match corresponding entries in `ref`. Here we'll do both.

```{r}
ne_dict <- data.frame(
  value = "NJ",
  replacement = "New Jersey",
  variable = "adm1"
)

(raw_match5 <- hmatch_settle(raw_remain4, ne_ref, dict = ne_dict,
                             fuzzy = TRUE, type = "resolve_inner"))

(raw_remain5 <- hmatch_settle(raw_remain4, ne_ref, dict = ne_dict,
                              fuzzy = TRUE, type = "resolve_anti"))
```
